---
title: "Hands-on Exercise 10"
subtitle: "Modelling Geographical Accesibility"
date: "20 Mar 2023"
date-modified: "`r Sys.Date()`"
number-sections: true
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

# Setting Up

```{r}
pacman::p_load(tmap, SpatialAcc, sf, 
               ggstatsplot, reshape2,
               tidyverse)
```

## Geospatial Data Wrangling

### Importing Data

```{r}
mpsz <- st_read(
  dsn = "data/geospatial",
  layer = "MP14_SUBZONE_NO_SEA_PL"
)
```

```{r}
hexagons <- st_read(
  dsn = "data/geospatial",
  layer = "hexagons"
)
```

```{r}
eldercare <- st_read(
  dsn = "data/geospatial",
  layer = "ELDERCARE"
)
```

### Updating CRS Information

```{r}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

### Cleaning and Updating Attribute Fields

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

## Aspatial Data Handling and Wrangling

### Importing Distance Matrix

```{r}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

Transform O-D matrix from thin to fat format.

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

Convert unit f measurement from meter to kilometer.

```{r}
distmat_km <- as.matrix(distmat/1000)
```

# Hansen Method for Modelling and Visualising Accessibility

## Computing Hansen's Accessibility

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

Rename the default name to `accHansen`

```{r}
colnames(acc_Hansen) <- "accHansen"
```

Convert data table into tibble format

```{r}
acc_Hansen <- as_tibble(acc_Hansen)
```

Combine tibble data frame with hexagon simple feature data frame

```{r}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## Visualising Hansen's Accessibility

### Extracting Map Extend

Extract extend of hexagons:

```{r}
mapex <- st_bbox(hexagons)
```

```{r}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

### Statistical Graphic Visualisation

```{r}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

Plot the distribution

```{r}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# KD2SFCA Method for Modelling and Visualising Accessibility

## Computing KD2SFCA's Accessibility

```{r}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))

colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- as_tibble(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## Visualising KD2SFCA's Accessibility

```{r}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical Graphic Visualisation

```{r}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

Plot the distribution

```{r}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# SAM (Spatial Accesibility Measure) Method for Modelling and Visualising Accessibility

## Computing SAM Accessibility

```{r}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- as_tibble(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## Visualising SAM's Accessibility

```{r}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical Graphic Visualisation

```{r}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

```{r}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```
